An Empirical Analysis of the Role of Energy in Economic Growth

### 2014-06-19 Article/EconGrowth2.Rnw
Line break changes. Made discussion of fitting processes more accurate.

In the context of linear models (regression), resamples are generally created
by residual resampling.  In our case, we formed resamples by adding to the fitted 
response (GDP) the product of a residual from the original model fit and 
and random sign ($-1$ or $1$, each with probability $0.5$).  For the case of 
Cobb-Douglas and LINEX models, this residual resampling occurs on the 
log-transformed data.

Intuitively, this method assumes that the residuals are indicative of 
the variability (from many potential sources) inherent in the data so that
it would be unsurprising if the residual from 
any particular year had been observed in a different year. 
Thus, a resampled response $\tilde Y$ can be computed as 
\[
\tilde Y_i' = \hat y_i \pm r_j
\]
where 
\[
r_i = y_i - \hat y_i
\]
both the sign ($\pm$) and the index of the residual ($j$, typically different from $i$) 
are chosen at random (with replacement). 
Resampling in this way has the consequence 
that a resampled time series could have non-unity first-year GDP ($\tilde Y_1 \neq 1$),
so a renormalization takes place using the initial GDP from the
resampled data as the new baseline, not the observed GDP in year 1.

\begin{equation} \label{eq:resample_y}
	\tilde y_i  = \frac{\tilde Y_i'}{\tilde Y_1'}
\end{equation}
to obtain a resampled indexed GDP time series.
We repeated the resampling process \Sexpr{numResamples()} times for each
combination of growth model and country, both with and without energy when appropriate.

The coefficients from the fit to a resampled time series
(the ``resample coefficients'') will be different from the coefficients 
obtained from the fit to historical data (the ``base coefficients'').
When these resample coefficients are highly variable, it is an indication
that the data do not determine the parameter estimates very precisely.
Lack of imprecision can stem from a number of factors, most obviously
a poor model fit, low model sensitivity to one or more parameters, 
or correlations between parameter estimates.
\authNote{This is not precisely true:
Large residuals in the fit to historical data (the ``base fit'')
are expected to produce large variation in resample coefficient values
(when the resampling is repeated \Sexpr{numResamples()} times),
indicating less certainty in the value
of those coefficients. 
Similarly, small base fit residuals are expected to produce small
variation in resample coefficient values, indicating more-precise knowledge
of the model coefficients.
--rjp}

>>>>

In the context of linear models (regression), resamples are generally created
by residual resampling.  In our case, we formed resamples by adding to the fitted 
response ($\hat{y}$) the product of a residual from the original model fit and 
random sign ($-1$ or $1$, each with probability $0.5$).  
For the case of Cobb-Douglas and LINEX models, 
this residual resampling occurs on the 
log-transformed data.

Intuitively, this method assumes that the residuals are indicative of 
the variability (from many potential sources) inherent in the data so that
it would be unsurprising if the residual from 
any particular year had been observed in a different year. 
Thus, a resampled response $\tilde Y'$ can be computed as 
\[
\tilde Y_i' = \hat y_i \pm r_j
\]
where 
\[
r_i = y_i - \hat y_i
\]
both the sign ($\pm$) and the index of the residual ($j$, typically different from $i$) 
are chosen at random (with replacement). 
% Resampling in this way has the consequence 
% that a resampled time series could have non-unity first-year GDP ($\tilde Y_1 \neq 1$),
% so a renormalization takes place using the initial GDP from the
% resampled data as the new baseline, not the observed GDP in year 1.
% \begin{equation} \label{eq:resample_y}
% 	\tilde y_i  = \frac{\tilde Y_i'}{\tilde Y_1'}
% \end{equation}
% to obtain a resampled indexed GDP time series.
We repeated the resampling process 1000 times for each
combination of growth model and country, 
both with and without energy as appropriate.

The coefficients from the fit to a resampled time series
(the ``resample coefficients'') will be different from the coefficients 
obtained from the fit to historical data (the ``base coefficients'').
When these resample coefficients are highly variable, 
it is an indication
that the data do not determine the parameter estimates very precisely.
Lack of precision can stem from a number of factors, most obviously
a poor model fit, low model sensitivity to one or more parameters, 
or correlations between parameter estimates.
\authNote{This is not precisely true:
Large residuals in the fit to historical data (the ``base fit'')
are expected to produce large variation in resample coefficient values
(when the resampling is repeated 1000 times),
indicating less certainty in the value
of those coefficients. 
Similarly, small base fit residuals are expected to produce small
variation in resample coefficient values, indicating more-precise knowledge
of the model coefficients.
--rjp}
